---
title: "Social_Distance_Visualization"
output: html_document
date: "2023-01-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(readxl)
library(stringr)
library(R.utils)
library(correlation)
library(corrplot)
library(ggpattern)
library(svglite)
library(tidytext)
colors <-c("#005f73","#0A9396","#E9D8A6","#ee9b00","#CA6702","#BB3E03")
```

```{r}
# Read in preprocessed data from each model 
XLNET_results <- read_csv("SD-results-XLNET-large.csv") %>% select(-1)
Distilbert_results <- read_csv("SD-results-DistilBERT.csv") %>% select(-1)
RB_results <- read_csv("SD-results-RoBERTa-base.csv") %>% select(-1)
RL_results <- read_csv("SD-results-RoBERTa-large.csv") %>% select(-1)
BTB_results <- read_csv("SD-results-BERTweet-base.csv") %>% select(-1)
BTL_results <- read_csv("SD-results-BERTweet-large.csv")%>% select(-1)
```



```{r}
##Merge results from each model 
all_results <- rbind(XLNET_results,Distilbert_results,RB_results,RL_results,BTB_results,BTL_results)
all_results <- all_results %>% 
  mutate(pachankis_conditions = ifelse(condition_categories=="Stigmatized",pachankis_conditions,condition))


## Assign category for each condition 
stigma_conditions <- read_csv("stigma_conditions.csv") %>% select(Pachankis_condition,Final_Category) %>% unique()
non_stigma_conditions <- read_csv("non_stigma_conditions.csv") %>% select(Condition, Final_Category) %>% unique()

# standardize names for Psoriasis condition
stigma_conditions  <- stigma_conditions %>%  mutate(Pachankis_condition =ifelse(str_detect(Pachankis_condition ,"Psoriasis") & str_detect(Pachankis_condition ,"current"),"Psoriasis (current;average severity)",Pachankis_condition ))
stigma_conditions  <-  stigma_conditions %>%  mutate(Pachankis_condition =ifelse(str_detect(Pachankis_condition ,"Psoriasis") & str_detect(Pachankis_condition ,"remitted"),"Psoriasis (remitted;average severity)",Pachankis_condition ))


stigma_conditions <- stigma_conditions %>% 
  mutate(condition = capitalize(Pachankis_condition)) %>% 
  select(condition,Final_Category)

non_stigma_conditions<- non_stigma_conditions %>% 
  mutate(condition = Condition) %>% select(condition,Final_Category)


conditions <- rbind(stigma_conditions,non_stigma_conditions)
all_results <- all_results %>% 
  left_join(conditions, by = c("pachankis_conditions"="condition")) %>% 
  mutate(category = Final_Category) %>% 
  select(-Final_Category)
```

# Evaluate the maximum probability and the minimum probability of the 50th token from all models' prediction 
```{r}
all_results %>% 
  group_by(prompt,Model,condition) %>% 
  summarise(min_prob = min(probs)) %>% 
  ungroup() %>% 
  summarise(max_minprob = max(min_prob),
            min_minprob = min(min_prob))
```

#Histogram of probability for each word 
```{r}

sum_probability_data <- all_results %>% 
  group_by(Model,prompt) %>%  # group by each model and each prompt 
  summarise(total_prediction_prob = sum(probs))  # sum the probability of all top 50 tokens in each prompt

all_results %>% 
  group_by(Model,prompt) %>% 
  summarise(total_prediction_prob = sum(probs)) %>% 
  ungroup() %>%  
  summarise(avg_summed_prob = mean(total_prediction_prob))

# svglite("summed_probability_for_top50_tokens_generated.svg",width = 10,height = 5)
sum_probability_data %>% 
  ggplot()+
  geom_histogram(aes(x = total_prediction_prob, fill=Model ))+
  scale_fill_manual(values =colors )+
  labs(x = "Summed Probability of Top 50 Tokens Per Prompt")+
  facet_wrap(~Model)+theme_bw()
# dev.off()
```

##Visualize Aggregated Results 
```{r}
####Read in files 
all_merged_condition_prob_XLNET<-read_csv("aggregated_SD_results_XLNET.csv") %>% mutate(Model="XLNet-large")
all_merged_condition_prob_RB<-read_csv("aggregated_SD_results_RB.csv") %>% mutate(Model="RoBERTa-base")
all_merged_condition_prob_RL<-read_csv("aggregated_SD_results_RL.csv") %>%  mutate(Model="RoBERTa-large")
all_merged_condition_prob_DB <- read_csv("aggregated_SD_results_DB.csv") %>% mutate(Model ="DistilBERT")
all_merged_condition_prob_BTB <- read_csv("aggregated_SD_results_BERTweet-Base.csv") %>% mutate(Model ="BERTweet-base")
all_merged_condition_prob_BTL <- read_csv("aggregated_SD_results_BERTweet-large.csv") %>% mutate(Model ="BERTweet-large")



###Merge all models data
all_data <- rbind(all_merged_condition_prob_DB,
                  all_merged_condition_prob_XLNET,
                  all_merged_condition_prob_RB,
                  all_merged_condition_prob_RL,
                  all_merged_condition_prob_BTB,
                  all_merged_condition_prob_BTL)

all_data <- all_data %>% mutate(condition =ifelse(str_detect(condition,"Psoriasis") & str_detect(condition,"current"),"Psoriasis (current;average severity)",condition))
all_data <- all_data %>% mutate(condition =ifelse(str_detect(condition,"Psoriasis") & str_detect(condition,"remitted"),"Psoriasis (remitted;average severity)",condition))

### Standardize condition categories 

non_stigma_conditions2 <- read_csv("non_stigma_conditions.csv") %>% select(Condition_StandardNaming, Final_Category)
colnames(non_stigma_conditions2)[1] <- "condition"
conditions <- rbind(stigma_conditions,non_stigma_conditions2)

all_data<- all_data %>% 
  left_join(conditions, by = "condition") %>% 
  mutate(category = Final_Category) %>% 
  select(-Final_Category) 
all_data <- all_data %>% mutate(category = ifelse(str_detect(category,"Diseases"),"Disease",category),
                                condition = ifelse(str_detect(condition,"Illteracy"),"Illiteracy",condition))



##
all_data_withoutBase <- all_data %>% filter(!condition=="None") %>% select(-1)
colnames(all_data_withoutBase)[4] <- "Condition Category"

```

# Visualize overall probability of a negative attitude for each condition (stigmatized and non-stigmatized)
```{r}
# list categories
category1 = c("Religion","Ethinicity","Sexuality","Disease","Mental Illness","Drug Use","Physical Traits","Disability","Socioeconomic Status","Education","Profession")

condition_overall_neg_prob_across_models <- all_data_withoutBase %>%
  filter(Label =="NEG") %>% 
  group_by(condition,category,`Condition Category`) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) 

# separate stigmatized and non-stigmatized condition data for visualization 
stigmatized_overall <- condition_overall_neg_prob_across_models %>% filter(`Condition Category`=="Stigmatized")
non_stigmatized_overall <- condition_overall_neg_prob_across_models %>% filter(`Condition Category`=="Non-stigmatized")

# save plot and visualize overall negative probability 
sd_visual_category1 <- all_data_withoutBase %>%
  filter(Label =="NEG") %>%  # look at negative probability 
  group_by(condition,category,`Condition Category`) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>% 
  filter(category %in% category1) %>% 
  ggplot()+
  geom_point(aes(x=category,y = condition_overall_neg_prob, color = `Condition Category`,
                 shape=`Condition Category`),
             size = 2)+
  scale_color_manual(values = c("#BB3E03","#005f73"))+
  theme_bw()+
   theme(axis.text.x = element_text(size = 13),
        legend.text = element_text(size = 10,lineheight = 2),
        legend.title = element_text(size = 10),
        legend.position = "bottom",
        strip.text.x = element_text(size = 10))+
  labs( y = "Overall Probability of Negative Attitude", x = "Conditions(in Category)")+ 
  ylim(0.25,1)+
  coord_flip()

# show plot
sd_visual_category1

# save plot for conditions that are not categorized 
sd_visual_category2 <- all_data_withoutBase %>%
  filter(Label =="NEG") %>% 
  group_by(condition,category,`Condition Category`) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>% 
  filter(!category %in% category1) %>% 
  ggplot()+
  geom_point(aes(x=category,y = condition_overall_neg_prob, color = `Condition Category`,shape=`Condition Category`),
             size = 2.5,,show.legend = FALSE)+
  scale_color_manual(values = c("#BB3E03","#005f73"))+
  theme_bw()+
     theme(axis.text.x = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        strip.text.x = element_text(size = 13))+
   labs( y = "Overall Probability of Negative Attitude", x = "Conditions")+
  ylim(0.25,1)+
  coord_flip()
sd_visual_category2


# function to put two plots together 
cowplot::plot_grid(sd_visual_category1, sd_visual_category2)
sd_both_condition_plot <- cowplot::plot_grid(sd_visual_category1, sd_visual_category2)

# ggsave2("Social_distance_Category.svg",width = 11,height = 5.5)

```
```{r}
# count the total number of conditions that has an overall negative probability greater than 0.6
 all_data_withoutBase %>%
  filter(Label =="NEG") %>% 
  group_by(condition,category,`Condition Category`) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>%
  filter(condition_overall_neg_prob>0.6) %>% 
  group_by(`Condition Category`) %>% 
   count()

```



# Calculate the overall average difference of probability of negative words between stigmatized and non-stigmatized conditiosn across models
```{r} 
colors <-c("#005f73","#0A9396","#E9D8A6","#ee9b00","#CA6702","#BB3E03")

all_data %>% 
  filter(Label=="NEG") %>% 
  group_by(Model,prompt_templates,condition_categories) %>% 
  summarise(avg_attitude = round(mean(condition_probs),digits = 2)) %>% 
  pivot_wider(names_from = condition_categories,values_from = avg_attitude) %>% 
  mutate(difference_of_negative_probs = Stigmatized - `Non-stigmatized`) %>% 
  ungroup() %>% 
  summarise(mean_diff = mean(difference_of_negative_probs))
```
```{r}
# average probability of negative words for each model and for each type of condition 
all_data %>% 
  filter(Label=="NEG") %>% 
  group_by(Model,prompt_templates,condition_categories) %>% 
  summarise(avg_attitude = round(mean(condition_probs),digits = 2)) %>% 
  pivot_wider(names_from = condition_categories,values_from = avg_attitude) %>% 
  mutate(Difference = Stigmatized-`Non-stigmatized`) %>% 
  ungroup() %>% 
  group_by(Model) %>% 
  summarise(avg_diff = mean(Difference))


```
```{r}
# overall difference between stigmatized conditions and nonstigmatized conditions by aggregating information from all templates and all models 
all_data %>% 
  filter(Label=="NEG") %>% 
  group_by(Model,prompt_templates,condition_categories) %>% 
  summarise(avg_attitude = round(mean(condition_probs),digits = 2)) %>% 
  pivot_wider(names_from = condition_categories,values_from = avg_attitude) %>% 
  mutate(difference_of_negative_probs = Stigmatized - `Non-stigmatized`) %>% 
  ungroup() %>% 
  group_by(Model) %>% 
  summarise(mean_diff = round(mean(difference_of_negative_probs),digits = 2)) %>% summarise(round(mean(mean_diff),digits=2))
```


# Identify conditions with an overall probability of negative attitude greater than 0.5
```{r}
# stigmatized conditions
all_data %>% 
  filter(Label =="NEG", condition_categories=="Stigmatized") %>% 
  group_by(condition) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>% 
  ungroup() %>% 
  filter(condition_overall_neg_prob >0.5)

# non-stigmatized conditions
all_data %>% 
  filter(Label =="NEG", condition_categories=="Non-stigmatized") %>% 
  group_by(condition) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>% 
  filter(condition_overall_neg_prob > 0.5)
```
# Visualize the overall probability of negative attitude for all conditions 
```{r}
all_data1 <- all_data
colnames(all_data1)[5]<- "Condition Categories"

# here we break the giant graph into two separate graphs 

# graph1 overall negative prob
# svglite("Aggregated_probability_of_negative_words_for_all_conditions_1.svg", width = 8, height = 11)
all_data1 %>% 

  filter(Label =="NEG") %>% 
  group_by(condition,`Condition Categories`) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>% 
  ungroup() %>% 
  filter(condition_overall_neg_prob >=0.5) %>%  # filter 
  ggplot()+
  geom_point(aes(x = reorder(condition,condition_overall_neg_prob), y = condition_overall_neg_prob, shape= `Condition Categories`,color = `Condition Categories`))+
  scale_color_manual(values = c("#BB3E03","#005f73"))+
  labs( x= "Condition",
        y = "Overall Probability of Negative Attitude")+
  theme(axis.text.x = element_text(size=12),
    legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))+
  ylim(c(0,1))+
  theme_bw()+
  coord_flip()
# dev.off()

#graph2.overall negative prob
# svglite("Aggregated_probability_of_negative_words_for_all_conditions_2.svg", width = 8, height = 8)
all_data1 %>% 
  filter(Label =="NEG") %>% 
  group_by(condition,`Condition Categories`) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) %>% 
  ungroup() %>% 
  filter(condition_overall_neg_prob <0.5) %>% 
  ggplot()+
  geom_point(aes(x = reorder(condition,condition_overall_neg_prob), y = condition_overall_neg_prob, shape= `Condition Categories`,color = `Condition Categories`))+
  scale_color_manual(values = c("#BB3E03","#005f73"))+
  labs( x= "Condition",
        y = "Overall Probability of Negative Attitude")+
  theme(axis.text.x = element_text(size=12),
    legend.text = element_text(size = 15),
        legend.title = element_text(size = 15))+
  ylim(c(0,1))+
  theme_bw()+
  coord_flip()

# dev.off()
```

# Generate figure for social distance results 
##Visualize baseline and condition graph together *attitude_prob_by_template_by_model is the baseline result that comes from the social_distance_base_line_analysis
```{r}
attitude_prob_by_template_by_model <- read_csv("aggregated_baseline_results_SD.csv")
condition_data <- all_data_withoutBase %>% filter(Label=="NEG") 
# svglite("baseline_and_conditions_SD.svg",width=13,height=5)
# png("baseline_and_conditions_SD.png")
ggplot()+
  geom_boxplot(data = attitude_prob_by_template_by_model,aes(x = prompt_templates,y = condition_probs),color ="#0A81D1")+
    scale_color_manual( values = c("Baseline Prompts" = "#E69F00")) +
  geom_boxplot_pattern(data= condition_data, aes(x = prompt_templates,
             y=condition_probs,
               fill=`Condition Category`,
               pattern = `Condition Category`),
             pattern_fill="black",
              color ="black",
              pattern_key_scale_factor = 0.5)+
  scale_pattern_manual(values = c( "none","stripe")) + 

  labs(col= "Conditions",
       y = "Probability of Negative Attitude",
       x = "Prompt templates")+
  ylim(c(0,1))+
  facet_grid(~Model)+
  theme_bw()+
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1,size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.position = "bottom",
        strip.text.x = element_text(size = 13))
# dev.off()
```

##Correlation between SD and Sentiment (MLMs and Downstream Sentiment Classification correlation!) 
```{r}
# read in the aggregated sentiment classification data 
negative_percentage_probability <-
  read_csv("aggregated_negative_percentage_classification_results.csv")

# summarise overall negative prob for each condition across all templates and all models
sd_overall_prob <- all_data %>% 
  filter(Label =="NEG") %>% 
  group_by(condition,condition_categories) %>% 
  summarise(condition_overall_neg_prob = mean(condition_probs)) 

# joined both datasets 
joint_sd_sentiment_data <- sd_overall_prob %>%
  left_join(negative_percentage_probability,by=c("condition"="Condition")) 

# remove a duplicate row 
joint_sd_sentiment_data<- joint_sd_sentiment_data[-64,] 

# make sure there is no more duplicate left
joint_sd_sentiment_stat<- joint_sd_sentiment_data %>% 
  ungroup() %>%
  select(condition_overall_neg_prob,Negative_Sentiment_Percentage) %>% 
  unique()

# calculate correlation 
correlation(joint_sd_sentiment_stat)
```
# detailed breakdown for condition in each model 
```{r}
all_data %>% 
  filter(Label =="NEG", condition_categories=="Stigmatized") %>% 
  group_by(Model,condition) %>% 
  summarise(avg_negative_prob = mean(condition_probs)) %>% 
  ggplot()+
  geom_point(aes(x = condition, y = avg_negative_prob, color = Model))+
  scale_color_manual(values = colors)+
  coord_flip()

all_data %>% filter(Label =="NEG", condition_categories=="Non-stigmatized") %>% group_by(Model,condition) %>% summarise(avg_negative_prob = mean(condition_probs)) %>% 
  ggplot()+
  geom_point(aes(x = condition, y = avg_negative_prob, color = Model))+
  scale_color_manual(values = colors)+
  coord_flip()

```


