---
title: "Sentiment Analysis"
output: html_document 
date: "2023-01-17"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE}
library(tidyverse)
library(R.utils)
library(ggpattern)
stigma_conditions <- read_csv("/Users/katelynmei/Documents/Github/DRG-Fall2022/stigma_conditions.csv") %>% select(Pachankis_condition,Final_Category) %>% unique()

sentiment_SIEBERT_stigma <- read_csv("SiEBERT_stigma_sentiment.csv") 
sentiment_SIEBERT_nonstigma <- read_csv("SiEBERT_nonstigma_sentiment.csv")
sentiment_TwitterRB_stigma <- read_csv("TwitterRB_stigma_sentiment.csv")
sentiment_TwitterRB_nonstigma <- read_csv("TwitterRB_nonstigma_sentiment.csv")
sentiment_BERTweet_nonstigma = read_csv("bertweet-base_nonstigma_sentiment.csv")
sentiment_BERTweet_stigma = read_csv("bertweet-base_stigma_sentiment.csv")

distil_bert_nonstigma <- read_csv("distilbert_nonstigma_sentiment.csv")
distil_bert_stigma <- read_csv("distilbert_stigma_sentiment.csv")

patterns2 <-c( "stripe","circle")
patterns3 <- c( "stripe","none","circle")
colors_2 <-c("#BB3E03","#457b9d")
colors_3 <-c("#BB3E03","#fdf0d5","#457b9d")
colors_4 <- c("#D8464B","#337064")
colors <-c("#005f73","#0A9396","#E9D8A6","#ee9b00","#CA6702","#BB3E03")

```

# Add in information for prompt category 
```{r}
add_information <- function(dataset,model_name,condition_name){
  dataset <- dataset %>% mutate(model = model_name,
                                condition_categories = condition_name,
                                prompt_templates =ifelse(str_detect(prompts,"They are"),
                                                         "they are",
                                                         ifelse(str_detect(prompts,"These are"),"these are","we are")))
  return(dataset)
}

SIEBERT_stigma_result <- add_information(sentiment_SIEBERT_stigma,"SiEBERT","Stigmatized")
SIEBERT_nonstigma_result <- add_information(sentiment_SIEBERT_nonstigma,"SiEBERT","Non-stigmatized")
SIEBERT_result <-rbind(SIEBERT_stigma_result,SIEBERT_nonstigma_result) %>% select(-1)

TwitterRB_stigma_result <- add_information(sentiment_TwitterRB_stigma,"TwitterRB_latest","Stigmatized")
TwitterRB_nonstigma_result <- add_information(sentiment_TwitterRB_nonstigma,"TwitterRB_latest","Non-stigmatized")
TwitterRB_result <- rbind(TwitterRB_stigma_result,TwitterRB_nonstigma_result) %>% select(-1)

BERTweet_stigma_result <- add_information(sentiment_BERTweet_stigma,"BERTweet","Stigmatized")
BERTweet_nonstigma_result <- add_information(sentiment_BERTweet_nonstigma,"BERTweet","Non-stigmatized")
BERTweet_result <- rbind(BERTweet_stigma_result,BERTweet_nonstigma_result) %>% select(-1)


distil_bert_nonstigma_sentiment<-  add_information(distil_bert_nonstigma,"DistilBERT base uncased finetuned SST-2","Non-stigmatized") %>% select(-1)
distil_bert_stigma_sentiment<-  add_information(distil_bert_stigma,"DistilBERT base uncased finetuned SST-2","Stigmatized") %>% select(-1)
distilbert_sentiment_result = rbind(distil_bert_nonstigma_sentiment,distil_bert_stigma_sentiment)

```


```{r}
#Function for data processing
aggregate_sentiment <- function(sentiment_data){
  #aggregate mean score of sentiment for condition
  aggregate_data <- sentiment_data %>%
    mutate(pachankis_conditions = ifelse(condition_categories=="Non-stigmatized",
                                         condition,
                                         pachankis_conditions))
  return(aggregate_data)
} 

distilbert_sentiment_result_aggregated <- distilbert_sentiment_result %>% aggregate_sentiment()
SIEBERT_result_aggregated <- SIEBERT_result %>% aggregate_sentiment()
TwitterRB_result_aggregated <- TwitterRB_result %>% aggregate_sentiment()
BERTweet_result_aggregated <- BERTweet_result %>% aggregate_sentiment()
```


#Merge all results from the four models and standardize spelling for conditions
```{r}

#Join results from all models
merged_sentiment_data <- rbind(TwitterRB_result_aggregated,
                              SIEBERT_result_aggregated,
                              BERTweet_result_aggregated,
                              distilbert_sentiment_result_aggregated) 

# Standardize sentiment words
merged_sentiment_data<- merged_sentiment_data %>% 
  mutate(Sentiment =ifelse(sentiment=="NEG"|sentiment=="NEGATIVE"|sentiment=="negative",
                           "Negative",
                      ifelse(sentiment=="NEU"|sentiment=="neutral","Neutral","Positive")) ) %>%
  select(-sentiment)

# Standardize naming for nonstigma conditions
nonstigmatized_conditions <- read_csv("/Users/katelynmei/Documents/Github/DRG-Fall2022/non_stigma_conditions.csv")
nonstigmatized_conditions_naming <- nonstigmatized_conditions %>% select(Condition,Condition_StandardNaming)

merged_sentiment_data <- merged_sentiment_data %>%
  left_join(nonstigmatized_conditions_naming,by = c("pachankis_conditions"="Condition")) %>%
  mutate(Condition = ifelse(condition_categories=="Stigmatized",
                            capitalize(pachankis_conditions),
                            Condition_StandardNaming) ) %>%
  select(-Condition_StandardNaming)

######## Fix categories 
## Assign category for each condition 
non_stigma_conditions2 <- nonstigmatized_conditions %>%
  select(Condition_StandardNaming,Final_Category)
colnames(non_stigma_conditions2)[1]<- "condition"

# conditions <- rbind(stigma_conditions,non_stigma_conditions2)

## filter out we are
merged_sentiment_data <- merged_sentiment_data %>% 
  filter(!prompt_templates=="we are") %>% 
  mutate(condition=ifelse(str_detect(condition,"Illteracy"),"Illiteracy",condition),
         Condition=ifelse(str_detect(Condition,"Illteracy"),"Illiteracy",Condition))


```

##Visualization starts here 

#Compare prompt templates difference
```{r}
# svglite("Prompt_templates_difference_SentimentAnalysis.svg", width = 5, height = 3)
merged_sentiment_data %>%
  group_by(Sentiment,prompt_templates) %>%
  count() %>%
  group_by(prompt_templates) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>%
  ggplot(aes(x = prompt_templates, y = percent_of_sentiment, fill = Sentiment))+
  geom_col()+
     geom_col_pattern(stat = "identity",
                   pattern_color = "white",
                   pattern_fill = "black",
                   pattern_key_scale_factor = 0.7,
                   aes(pattern = Sentiment))+
   scale_pattern_manual(values =patterns3) +
  scale_fill_manual(values =colors_3)+
  labs(y ="Distribution of Sentiment",
       x = "Prompt Templates")+
  theme_bw()
 # dev.off()
# svglite("Prompt_templates_difference_byModel_SentimentAnalysis.svg", width = 8, height = 3)
merged_sentiment_data %>%
  filter(!prompt_templates=="we are") %>% 
  group_by(model,Sentiment,prompt_templates) %>%
  count() %>%
  group_by(model, prompt_templates) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>%
  ggplot(aes(x = prompt_templates, y = percent_of_sentiment, fill = Sentiment))+
  geom_col()+
     geom_col_pattern(stat = "identity",
                   pattern_color = "white",
                   pattern_fill = "black",
                   pattern_key_scale_factor = 0.7,
                   aes(pattern = Sentiment))+
   scale_pattern_manual(values =patterns3) +
  scale_fill_manual(values =colors_3)+
  labs(y ="Distribution of Sentiment",
       x = "Prompt Templates")+
  facet_grid(~model,labeller = label_wrap_gen(width=15))+
  theme_bw()+
      theme(axis.text.x = element_text(size = 10),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 10),
        legend.position = "bottom",
        strip.text.x = element_text(size = 7))
# dev.off()

# Evaluate difference between templates in each model 
 merged_sentiment_data %>%
  group_by(model,Sentiment,prompt_templates) %>%
  count() %>%
  group_by(model, prompt_templates) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>% 
   select(-n) %>% 
   pivot_wider(names_from = prompt_templates, values_from = percent_of_sentiment) %>% 
   filter(Sentiment=="Negative") %>% 
   mutate(diff = round(`these are`-`they are`,digits = 2))
```

```{r Visualize ration of sentiment}
# svglite("All_models_between_conditions_SentimentAnalysis.svg", width = 13, height = 5)
merged_sentiment_data %>%
  group_by(Sentiment,condition_categories,model) %>%
  count() %>%
  group_by(condition_categories,model) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>%
     ggplot(aes(x = condition_categories, y = percent_of_sentiment, fill = Sentiment))+
  geom_col()+
   geom_col_pattern(stat = "identity",
                   pattern_color = "white",
                   pattern_fill = "black",
                   pattern_key_scale_factor = 0.7,
                   aes(pattern = Sentiment))+
  scale_pattern_manual(values =patterns3) +
  scale_fill_brewer(palette = "RdBu")+
  labs( y = "Sentiment Classification", x = "Condition Categories")+
  scale_fill_manual(values =colors_3)+
  theme_bw()+
    theme(axis.text.x = element_text(size = 10),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.position = "bottom",
        strip.text.x = element_text(size = 13))+
  facet_grid(~model,labeller = label_wrap_gen(width=13))
# dev.off()

###Graphs for they are 
# svglite("All_models_between_conditions_Theyare_SentimentAnalysis.svg", width = 13, height = 5)
merged_sentiment_data %>%
  filter(prompt_templates=="they are") %>%
  group_by(Sentiment,condition_categories,model) %>%
  count() %>%
  group_by(condition_categories,model) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>%
     ggplot(aes(x = condition_categories, y = percent_of_sentiment, fill = Sentiment))+
   geom_col_pattern(stat = "identity",
                   pattern_color = "white",
                   pattern_fill = "black",
                   pattern_key_scale_factor = 0.7,
                   aes(pattern = Sentiment))+
  scale_pattern_manual(values =patterns3) +
  labs( y = "Sentiment Classification", x = "Condition Categories",title="Prompt Templates: 'They are'")+
  scale_fill_manual(values =colors_3)+
  theme_bw()+
    theme(axis.text.x = element_text(size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.position = "bottom",
        strip.text.x = element_text(size = 13))+
  facet_grid(~model,labeller = label_wrap_gen(width=15))
# dev.off()

###graphs for these are
# svglite("All_models_between_conditions_Theseare_SentimentAnalysis.svg", width = 13, height = 5)
merged_sentiment_data %>%
  filter(prompt_templates=="these are") %>%
  group_by(Sentiment,condition_categories,model) %>%
  count() %>%
  group_by(condition_categories,model) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>%
     ggplot(aes(x = condition_categories, y = percent_of_sentiment, fill = Sentiment))+
   geom_col_pattern(stat = "identity",
                   pattern_color = "white",
                   pattern_fill = "black",
                   pattern_key_scale_factor = 0.7,
                   aes(pattern = Sentiment))+
  scale_pattern_manual(values =patterns3) +
  labs( y = "Sentiment Classification", x = "Condition Categories",title="Prompt Templates: 'These are'")+
  scale_fill_manual(values =colors_3)+
  theme_bw()+
      theme(axis.text.x = element_text(vjust = 1, hjust=1,size = 13),
        legend.text = element_text(size = 13),
        legend.title = element_text(size = 13),
        legend.position = "bottom",
        strip.text.x = element_text(size = 13))+
  facet_grid(~model,labeller = label_wrap_gen(width=15))
# dev.off()
```


# In general if we take all sentiment classification account for each condition, what proportion of prompts will be classified as negative?
```{r}

colors_portion <-c("#457b9d","#BB3E03")
#"#98c1d9","#e29578"
colors_portion2 <-c("#BB3E03","#fdf0d5")

#### Regroup sentiment into two categories 
# break the giant graph into two graphs 

# graph1 
# svglite("Condition_Ranking_SentimentAnalysis_1.svg", width = 8, height = 9)
merged_sentiment_data %>% 
  mutate(New_sentiment = ifelse(Sentiment=="Negative","Negative","Non-negative")) %>% 
  filter(condition_categories=="Stigmatized",
         !pachankis_conditions%in%("a person")) %>% 
  group_by(Condition) %>% 
  count(New_sentiment) %>% 
  mutate(Sentiment_proportion = n/sum(n)) %>% 

  mutate(dummy_score = case_when(New_sentiment == "Negative" ~ Sentiment_proportion,
                          TRUE ~ 1 - Sentiment_proportion)) %>% # use this to reorder variable! 
  filter(dummy_score >0.5) %>% 
  ggplot()+
  geom_col(aes(x = reorder(Condition,dummy_score), y = Sentiment_proportion, fill = New_sentiment),show.legend = TRUE) +
  scale_fill_manual(values =c("#98c1d9","#e29578"))+
  labs(fill = "Sentiment", x = "Stigmatized Conditions", y = "Distribution of Sentiment")+
  theme_bw()+
  coord_flip()+
    ggtitle("Sentiment Classification Outcomes for Stigmatized Conditions")+
      theme(plot.title = element_text(size=10,hjust = 0.5))
# dev.off()

# graph2 
# svglite("Condition_Ranking_SentimentAnalysis_2.svg", width = 8, height = 7)
merged_sentiment_data %>% 
  mutate(New_sentiment = ifelse(Sentiment=="Negative",
                                "Negative",
                                "Non-negative")) %>% 
  filter(condition_categories=="Stigmatized",
         !pachankis_conditions%in%("a person")) %>% 
  group_by(Condition) %>% 
  count(New_sentiment) %>% 
  mutate(Sentiment_proportion = n/sum(n)) %>% 
  mutate(dummy_score = case_when(New_sentiment == "Negative" ~ Sentiment_proportion,
                          TRUE ~ 1 - Sentiment_proportion)) %>% # use this to reorder variable! 
  filter(dummy_score <=0.5) %>% 
  ggplot()+
  geom_col(aes(x = reorder(Condition,dummy_score), y = Sentiment_proportion, fill = New_sentiment),show.legend = TRUE) +
  scale_fill_manual(values =c("#98c1d9","#e29578"))+
  labs(fill = "Sentiment", x = "Stigmatized Conditions", y = "Distribution of Sentiment")+
  theme_bw()+
  coord_flip()+
  ggtitle("Sentiment Classification Outcomes for Stigmatized Conditions")+
    theme(plot.title = element_text(size=10,hjust = 0.5))
# dev.off()



# svglite("Condition_Non-Stigmatized_Ranking_SentimentAnalysis.svg", width = 8, height = 5)
merged_sentiment_data %>% 
  mutate(New_sentiment = ifelse(Sentiment=="Negative","Negative","Non-negative")) %>% 
  filter(condition_categories=="Non-stigmatized") %>% 
  group_by(Condition) %>% 
  count(New_sentiment) %>% 
  mutate(Sentiment_proportion = n/sum(n)) %>% 
  mutate(dummy_score = case_when(New_sentiment == "Negative" ~ Sentiment_proportion,
                          TRUE ~ 1 - Sentiment_proportion)) %>%
  ggplot()+
  geom_col(aes(x = reorder(Condition,dummy_score), y = Sentiment_proportion, fill = New_sentiment),show.legend = TRUE) +
  scale_fill_manual(values = c("#98c1d9","#e29578"))+
  labs( fill = "Sentiment",x = "Non-stigmatized Conditions", y = "Distribution of Sentiment")+
  theme_bw()+
  ggtitle("Sentiment Classification Outcomes for Non-stigmatized Conditions")+
    theme(plot.title = element_text(size=10,hjust = 0.5))+
  coord_flip()
# dev.off()


```
```{r}

# Calculate the proportion of stigmatized conditions are classifed with more than 50 percent negative sentiments 
merged_sentiment_data %>% 
  filter(condition_categories=="Stigmatized",
         !pachankis_conditions%in%("a person")) %>% 
  group_by(Condition) %>% 
  count(Sentiment) %>% 
  mutate(sentiment_proportion = n/sum(n)) %>% 
  filter(Sentiment=="Negative") %>% 
  mutate(greater_or_equal_50 = ifelse(sentiment_proportion >0.5,1,0)) %>%
  group_by(greater_or_equal_50) %>% 
  count() %>% ungroup() %>% mutate(percent_thatisGreaterorEqualTo_50 = n/93)

# check total conditions match what we have 
merged_sentiment_data %>% ungroup() %>% filter(condition_categories =="Stigmatized") %>% select(pachankis_conditions) %>% unique() # in total we have 91 conditions 
merged_sentiment_data %>% ungroup() %>% filter(!pachankis_conditions%in%("a person"),condition_categories =="Non-stigmatized") %>% select(Condition) %>% unique() # in total we have  

# Calculate the proportion of nonstigmatized conditions are classifed with more than 50 percent negative sentiments 
merged_sentiment_data %>% 
  filter(condition_categories=="Non-stigmatized",
         !pachankis_conditions%in%("a person")) %>% 
  group_by(Condition) %>% 
  count(Sentiment) %>% 
  mutate(sentiment_proportion = n/sum(n)) %>% 
  filter(Sentiment=="Negative") %>% 
  filter(sentiment_proportion >0.5)

```

## All models compare categories (output negative percentage probability for correlation analysis)
```{r}
condition_sentiment_percentage <- merged_sentiment_data %>%  
  mutate(Binary_sentiment =ifelse(Sentiment =="Negative","Negative","Non-negative")) %>% 
  group_by(Binary_sentiment,Condition,category) %>%
  count() %>%
  ungroup() %>% 
  group_by(Condition) %>%
  mutate(percent_of_sentiment = n/sum(n)) %>% 
  mutate(Negative_Sentiment_Percentage = ifelse(Binary_sentiment=="Negative",percent_of_sentiment,1-percent_of_sentiment))


negative_percentage_probability <- condition_sentiment_percentage %>% select(Condition,  Negative_Sentiment_Percentage) %>% unique()
negative_percentage_probability
# write_csv(negative_percentage_probability,"aggregated_negative_percentage_classification_results.csv")
condition_sentiment_percentage %>%
  ungroup() %>% 
  group_by(category,Binary_sentiment) %>% # calculate the percent of sentiment for each condition 
  summarise(mean_sentiment_prob = mean(percent_of_sentiment))
```



